Sparse Spectral Methods on Disk Slices

### 2019-02-08 posters/jamboree2019.tex
First draft of poster for HD. Minor edits to paper.

\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for rotated page geometry
\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{caption}
\usepackage{subcaption}
\usepackage{float}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{bbm}
\usepackage{mleftright}
\usepackage{todonotes}

\graphicspath{ {images/} }

@@

\usepackage{natbib}
\usepackage{url}
%\bibliographystyle{elsarticle-harv}
\bibliographystyle{plain}

@@

%\usepackage{media9}
%\graphicspath{ {../sphere/plots} }

\title{Sparse spectral and $p$-finite element methods on disk-slices and rhombi I: zero boundary conditions}
\author{Sheehan Olver, Ben Snowball}
%\date{}							% Activate to display a given date or no date

\begin{document}

\maketitle

\begin{abstract}
In recent years, sparse spectral methods have been derived using hierarchies of classical orthogonal polynomials on intervals, disks, and triangles. In this work we extend this methodology to a hierarchy of non-classical orthogonal polynomials on disk slices (e.g. a half disk) and rhombi. This builds on the observation that sparsity is guaranteed due to the boundary being defined by an algebraic curve, and that the entries of partial differential operators can be determined using formulae in terms of (non-classical) univariate orthogonal polynomials. 
\end{abstract}

%
\section{Introduction}

This paper develops sparse spectral methods for solving linear partial differential equations on a special class of geometries that includes disk slices and rhombi.  
More precisely, let  \(\rho \: : \: (\alpha,\beta) \: \to (0,\infty)\) be such that either:

\begin{enumerate}
\item  \(\rho\) is a degree 1 polynomial, or 
\item \(\rho\) is the square root of a non-negative degree \(\le\) 2 polynomial, \(-\gamma = \delta > 0\), and \(w_2\) is an even function.
\end{enumerate}

We consider the solution of partial differential equations on the domain
\begin{align}
\Omega := \{(x,y) \in \R^2 \quad | \quad \alpha < x < \beta, \: \gamma \rho(x) < y < \delta \rho(x)\}
\end{align}
For simplicity of discussion we focus on the half-disk, where $\rho(x) = \sqrt{1-x^2}$,  $(\alpha,\beta) = (0, 1)$, and  $(\gamma, \delta)  = (-1,1)$, with discussion of extension to other geometries in the appendix.

We show that partial differential equations become sparse linear systems when viewed as acting on expansions involving a family of orthogonal polynomials that  generalise Jacobi polynomials, mirroring the ultraspherical spectral method for ordinary differential equations \cite{SOATUltraspherical} and its analogue on the triangle \cite{SOATGVTriangle}.  On the half-disk the family of weights we consider are of the form
$$
W^{(a,b)}(x,y) = x^a (1-x-y)^b
$$
with corresponding OPs denoted $P_{n,k}^{(a,b)}(x,y)$. Sparsity comes from expanding the domain and range of an operator  using different polynomials. Where the sparsity pattern and entries derived in \cite{SOATGVTriangleRecurrence,SOATGVTriangle}  results from manipulations of Jacobi polynomials, in the present work we use a more general integration-by-parts argument to deduce the sparsity structure, alongside quadrature rules to determine the entries.  Furthermore, we use this framework to derive sparse $p$-finite element methods that are analogous to those of Beuchler and Sch\"oberl on tetrahedra \cite{TriangleFEM}.

The motivation for this work is solving partial differential equations on sub-domains of the sphere. In particular, OPs on a half-sphere can be represented using two families of OPs on the half-disk. Constructing sparse spectral methods for surface PDEs on half-spheres, spherical caps, and spherical triangles is future work, and has applications in weather prediction. Other extensions include a full $hp$-finite element method on sections of a disk, which has applications in fluids.

 We present our procedure  to gain a (two-parameter) family of 2D orthogonal polynomials (OPs) on the half disk domain, by combining 1D OPs to the interval, to form 2D OPs on the disk. We will derive a quadrature rule on the half disk that can be used to expand a function in the OP basis up to a given order. We will demonstrate that these families will have sparse operators (including Jacobi operators representing multiplication by $x$ and $y$, and partial differential operators), and how we can implement function evaluation using the coefficients of the expansion of a given function using the Clenshaw algorithm. We can extend this work directly by projecting up to 3D OPs on the hemisphere, where sparse operators, function evaluation and expansion etc. can be found similarly.

\section{Orthogonal polynomials on the half disk}

In this section we outline some basic properties of $P_{n,k}^{(a,b)}(x,y)$. The symmetry in the weight allows us to express the polynomials in terms of 1D OPs, and deduce certain properties such as recurrence relationships.

\subsection{Explicit construction}

@@

For the half disk, the weight $W^{(a,b)}(x,y)$ results from setting:

\begin{align}
\begin{cases}
(\alpha,\beta) &:= (0.1) \\
(\gamma,\delta) &:= (-1,1) \\
\rho(x) &:= (1-x^2)^{\half} \\
w_1(x) &:= x^a \: (1-x^2)^b \\
w_2(x) &:= (1-x^2)^b = (1-x)^b \: (1+x)^b,
\end{cases}
\end{align}

The weight $w_2(x)$ is an ultraspherical weight, and the corresponding OPs are   the Jacobi polynomials  \(\{P_n^{(b, b)}\}\). The weight $w_1(x)$ is non-classical and we introduce the notation \(\{H_n^{(a, b)}\}\) for the orthonormal polynomials. Thus we arrive at  a two parameter family of  2D orthogonal polynomials \(\{\Pnk\}\) given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
\begin{align}
 \Pnk^{(a,b)}(x,y) := H_{n-k}^{(a, b+k+\half)}(x) \: \rho(x)^k \: P_k^{(b,b)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
\end{align}
orthogonal on \(\Omega\) with respect to the weight $W^{(a,b)}(x,y)$.

\subsection{Jacobi matrices}

Express the three-term recurrences associated to $H_n^{(a,b)}$ and $P_n^{(b,b)}$ as
\begin{align}
x H_n^{(a,b)}(x) &= \beta_n^{(a,b)} H_{n+1}^{(a,b)}(x) + \alpha_n^{(a,b)} H_n^{(a,b)}(x) + \beta_{n-1}^{(a,b)} H_{n-1}^{(a,b)}(x) \\
y P_n^{(b,b)}(y) &= \delta_n^{(b)} P_{n+1}^{(b,b)}(y) + \gamma_n^{(b)} P_n^{(b,b)}(y) + \delta_{n-1}^{(b)} P_{n-1}^{(b,b)}(y),
\end{align}
where we note that \(\gamma_n^{(b)} = 0\) \(\forall n = 0,1,2,\dots\). We can use these to determine the 2D recurences for $P_{n,k}^{(a,b)}(x,y)$:

\begin{lemma}
$\Pnk^{(a,b)}(x,y)$ satisfy the following 3-term recurrences:
\begin{align}
x \Pnk^{(a,b)}(x,y) &= \alphaab_{n,k,1} \: P_{n-1, k}^{(a,b)}(x, y) + \alphaab_{n,k,2} \: P_{n, k}^{(a,b)}(x, y) + \alphaab_{n+1,k,1} \: P_{n+1, k}^{(a,b)}(x, y), \\
y \Pnk^{(a,b)}(x,y) &= \betaab_{n,k,1} \: P_{n-1, k-1}^{(a,b)}(x, y) + \betaab_{n,k,2} \: P_{n-1, k+1}^{(a,b)}(x, y) \nonumber \\
		& \quad \quad + \betaab_{n,k,3} \: P_{n, k-1}^{(a,b)}(x, y) + \betaab_{n,k,4} \: P_{n, k+1}^{(a,b)}(x, y) \nonumber \\
		& \quad \quad + \betaab_{n,k,5} \: P_{n+1, k-1}^{(a,b)}(x, y) + \betaab_{n,k,6} \: P_{n+1, k+1}^{(a,b)}(x, y),
\end{align}
for \((x,y) \in \Omega\), where
\begin{align}
\alphaab_{n,k,1} &:= \beta_{n-k-1}^{(a, b+k+\half)} \\
\alphaab_{n,k,2} &:= \alpha_{n-k}^{(a, b+k+\half)} \\
\betaab_{n,k,1} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n-1,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k}^{(a, b+k-\half)}>_{H^{(a, b+k+\half)}} \\
\betaab_{n,k,2} &:= \frac{\delta_{k}^{(b)}}{|| P_{n-1,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k-2}^{(a, b+k+\frac{3}{2})}>_{H^{(a, b+k+\frac{3}{2})}} \\
\betaab_{n,k,3} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k+1}^{(a, b+k-\half)}>_{H^{(a, b+k+\half)}} \\
\betaab_{n,k,4} &:= \frac{\delta_{k}^{(b)}}{|| P_{n,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k-1}^{(a, b+k+\frac{3}{2})}>_{H^{(a, b+k+\frac{3}{2})}} \\
\betaab_{n,k,5} &:= \frac{\delta_{k-1}^{(b)}}{|| P_{n+1,k-1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k+2}^{(a, b+k-\half)}>_{H^{(a, b+k+\half)}} \\
\betaab_{n,k,6} &:= \frac{\delta_{k}^{(b)}}{|| P_{n+1,k+1}^{(a,b)} ||_{W^{(a,b)}}^2} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{n-k}^{(a, b+k+\frac{3}{2})}>_{H^{(a, b+k+\frac{3}{2})}}. 
\end{align}
where $|| p(x,y) ||_{W^{(a,b)}}^2 := \int_\Omega p(x,y)^2 \: \Wab(x,y) \: d\mathbf{x}$ and $\ip<p(x), q(x)>_{H^{(a, b)}} := \int_\alpha^\beta p(x) \: q(x) \: x^a \: (1-x^2)^b \: dx$, i.e the 1D inner product using the $\{H^{(a,b)}_n\}$ family weight.

\end{lemma}

\begin{proof}
Note that \(\ip<y \Pnkab, P_{m,j}^{(a,b)}>_\Wab = 0\) for \(m < n-1\) (and for \(m > n+2\)). Thus for \(m = n-1, n, n+1,\) \(j = 0,\dots,m:\)
\begin{align}
\ip<y \Pnkab, P_{m,j}^{(a,b)}>_\Wab &= \int \int_\Omega \Pnkab(x,y) \: \Pmjab(x,y) \: y \: \Wab(x,y) \: dy \: dx \\
&= \Big( \int^1_0 H_{n-k}^{(a, b+k+\half)}(s) \: H_{m-j}^{(a, b+j+\half)}(s) \: s^a \: \rho(s)^{2b+k+j+2} \: ds \Big) \nonumber \\
& \quad \quad \quad\cdot \: \Big( \int^1_{-1} P_k^{(b,b)}(t) \: P_j^{(b,b)}(t) \: t \: (1-t^2)^b \: dt \Big) \\
&= \begin{cases}
    	\delta_k^{(b)} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{m-k-1}^{(a, b+k+\frac{3}{2})}>_{H^{(a, b+k+\frac{3}{2})}} \quad& \text{if } j = k+1 \\
	\delta_{k-1}^{(b)} \: \ip<H_{n-k}^{(a, b+k+\half)}, H_{m-k+1}^{(a, b+k-\half)}>_{H^{(a, b+k+\half)}} \quad& \text{if } j = k-1 \\
	0 & \text{otherwise}
      \end{cases}.
\end{align}

\end{proof}

Three-term recurrences lead to Jacobi operators that correspond to multiplication by $x$ and $y$. Define, for some $N \in \No$:
\begin{align}
\bigPab_n := \begin{bmatrix}
		P^{(a,b)}_{n,-n}(x,y) \\
		\vdots \\
		P^{(a,b)}_{n,n}(x,y)
	\end{bmatrix} \in \R^{n+1}, 
\quad \quad 
\bigPab := \begin{bmatrix}
		\bigPab_0 \\
		\bigPab_1 \\
		\vdots \\
		\bigPab_N
	\end{bmatrix}
\end{align}
and set \(J_x^{(a,b)}, J_y^{(a,b)}\) as the \(\half (N+1)(N+2) \times \half (N+1)(N+2)\) Jacobi matrices corresponding to
\begin{align}
J_x^{(a,b)} \bigPab(x,y) = x \bigPab(x,y), \quad J_y^{(a,b)} \bigPab(x,y) = y \bigPab(x,y).
\end{align}
Then  \(J_x^{(a,b)}, J_y^{(a,b)}\) have the Banded-Block-Banded structures
\begin{align}
J_{x/y}^{(a,b)} &= \begin{bmatrix}
		B^{x/y}_0 & A^{x/y}_0 & & & & \\
		C^{x/y}_1 & B^{x/y}_1 & A^{x/y}_1 & & & \\
		& C^{x/y}_2 & B^{x/y}_2 & A^{x/y}_2  & & & \\
		& & C^{x/y}_3 & \ddots & \ddots & \\
		& & & \ddots & \ddots & \ddots \\
	\end{bmatrix}
\end{align}
where
\begin{align}
A^x_n &:= \begin{bmatrix}
		\alphaab_{n+1,0,1} & 0 & \hdots & 0 \\
		& \ddots & & \vdots & \\
		& & \alphaab_{n+1,n,1} & 0 \\
	    \end{bmatrix} \in \R^{(n+1)\times(n+2)}, \quad n = 0,1,2,\dots,N-1 \\
B^x_n &:= \begin{bmatrix}
		\alphaab_{n,0,2} & & \\
		& \ddots & \\
		& & \alphaab_{n,n,2} \\
	    \end{bmatrix} \in \R^{(n+1)\times(n+1)} \quad n = 0,1,2,\dots,N \\
C^x_n &:= \big( A^x_n \big)^T \in \R^{(n+1)\times n},  \quad n = 1,2,\dots,N, \\ 
\nonumber \\
A^y_n &:= \begin{bmatrix}
		0 & \betaab_{n,0,6} & & & \\
		\betaab_{n,1,5} & 0 & & & \\
		& \ddots & \ddots & \ddots & \\
		& & \betaab_{n,n,5}& 0 & \betaab_{n,n,6} \\
	    \end{bmatrix} \in \R^{(n+1)\times(n+2)}, \quad n = 0,1,2,\dots,N-1 \\
B^y_n &:= \begin{bmatrix}
		0 & \betaab_{n,0,4} & & \\
		\betaab_{n,1,3} & 0 & \ddots & \\
		& \ddots & \ddots & \betaab_{n,n-1,4} \\
		& & \betaab_{n,n,3} & 0
	    \end{bmatrix} \in \R^{(n+1)\times(n+1)}  \quad n = 0,1,2,\dots,N \\
C^y_n &:= \begin{bmatrix}
		0 & \betaab_{n,0,2} & & \\
		\betaab_{n,1,1} & 0 & \ddots & \\
		& \ddots & \ddots & \betaab_{n,n-2,2} \\
		& & \ddots & 0 \\
		& & & \betaab_{n,n,1} \\
	    \end{bmatrix} \in \R^{(n+1)\times n}, \quad n = 1,2,\dots,N.
\end{align}

Note that the sparsity of the Jacobi matrices (in particular the sparsity of the sub-blocks) comes from the natural sparsity of the three-term recurrences of the 1D OPs. In other words, sparsity is not dependent on the domain.

\subsection{Building the OPs}

Combining each system in (71) we can write the block-tridiagonal system
\begin{align}
\renewcommand\arraystretch{1.3}
\mleft[
\begin{array}{cccc}
		1 & & & \\
		B_0-G_0(x,y) & A_0 & & \\
		C_1 & B_1-G_1(x,y) & \quad A_1 \quad & \\
		& C_2 & \ddots & \ddots \\
		& & \ddots &
\end{array}
\mright]
\bigPab(x,y)
=
\begin{bmatrix}
	P^{(a,b)}_{0,0} \\ 0 \\ 0 \\ \vdots \\ 0 \\
\end{bmatrix}.
\end{align}
where we note \(P^{(a,b)}_{0,0}(x,y) \equiv H_0^{(a,b)} \: P_0^{(b,b)}\), and for each \(n = 0,1,\dots,N\):
\begin{align}
A_n &:= \begin{bmatrix}
		A^x_n \\
		A^y_n \\
		A^z_n
	    \end{bmatrix} \in \R^{2(n+1)\times(n+3)} \quad (n \ne N), \quad
C_n := \begin{bmatrix}
		C^x_n \\
		C^y_n
	    \end{bmatrix} \in \R^{2(n+1)\times n} \quad (n \ne 0), \nonumber \\
B_n &:= \begin{bmatrix}
		B^x_n \\
		B^y_n
	    \end{bmatrix} \in \R^{2(n+1)\times(n+1)}, \quad
G_n(x,y,z) := \begin{bmatrix}
		xI_{n+1} \\
		yI_{n+1}
	    \end{bmatrix} \in \R^{2(n+1)\times(n+1)}.
\end{align}

For each \(n = 0,1,\dots,N-1\) let \(\Dnt\) be any matrix that is a left inverse of \(A_n\), i.e. such that \(\Dnt A_n = I_{n+3}\). Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the \(\Dnt\)'s, we obtain a lower triangular system \citep{dunkl2014orthogonal} (p78), which can be expanded to obtain the recurrence:
\begin{align}
\begin{cases}
\bigPab_{-1}(x,y) := 0 \\
\bigPab_{0}(x,y) := P^{(a,b)}_{0,0} \\
\bigPab_{n+1}(x,y) = -\Dnt (B_n-G_n(x,y)) \bigPab_n - \Dnt C_n  \,\bigPab_{n-1}(x,y), \quad n = 1,2,3,\dots,N.
\end{cases}
\end{align}

Note that we can define \(\Dnt\) as follows. For \(n\) even:
\begin{align}
\Dnt := \begin{bmatrix}
		\frac{1}{\alphaab_{n+1,0,1}} & & &  \\
		& \ddots & & & \\
		& & \frac{1}{\alphaab_{n+1,n,1}} & \\
		0 & \hdots & 0 & \eta_{m-1} & \hdots & 0 & \eta_1 & 0 & \eta_0
	    \end{bmatrix},
\end{align}
where
\begin{align}
m &= \frac{n}{2} + 1, \quad \eta_0 = \frac{1}{\betaab_{n,n,6}}, \nonumber \\
\eta_k &= -\frac{\betaab_{n,n-2(k-1),5} \: \eta_{k-1}}{\betaab_{n,n-2k,6}} \quad k = 1,\dots,m-1.
\end{align}

For \(n\) odd:

\begin{align}
\Dnt := \begin{bmatrix}
		\frac{1}{\alphaab_{n+1,0,1}} & &  \\
		& \ddots & & &  \\
		& & \ddots & & \\
		& & & \frac{1}{\alphaab_{n+1,n,1}} & \\
		\xi & 0 & \hdots & 0 & 0 & \eta_{m-2} & \hdots & 0 & \eta_0
	    \end{bmatrix},
\end{align}
 where
\begin{align}
m &= \frac{n+1}{2} + 1, \quad \eta_0 = \frac{1}{\betaab_{n,n,6}}, \nonumber \\
\eta_k &= -\frac{\betaab_{n,n-2(k-1),5} \: \eta_{k-1}}{\betaab_{n,n-2k,6}} \quad k = 1,\dots,m-2, \\
\xi &= -\frac{\betaab_{n,1,5} \: \eta_{m-2}}{\alphaab_{n+1,0,1}}. \nonumber
\end{align}

%
\section{Sparse partial differential operators}

\begin{figure} 
\center
\begin{tikzpicture} 
% Box/square (4x4)
\draw[black,solid,ultra thin] (0,0)--(0,4);
\draw[black,solid,ultra thin] (0,0)--(4,0);
\draw[black,solid,ultra thin] (4,0)--(4,4);
\draw[black,solid,ultra thin] (0,4)--(4,4);
% Dots
\draw[black,fill=black] (0,0) circle (.5ex);
\draw[black,fill=black] (2,2) circle (.5ex);
\draw[black,fill=black] (4,4) circle (.5ex);
\draw[black,fill=black] (0,2) circle (.5ex);
\draw[black,fill=black] (0,4) circle (.5ex);
% Arrows
\draw[black,thick,->] (0.2,0.2)--(1.8,1.8);
\draw[black,thick,->] (2.2,2.2)--(3.8,3.8);
\draw[black,thick,->] (0,0.2)--(0,1.8);
\draw[black,thick,->] (0,2.2)--(0,3.8);
\draw [black,thick,->] (0.2,4.2) to [out=20,in=160] (3.8,4.2);
% Node (parameter) labels
\draw[] (0,0) node[anchor=east] {$(0,0)$};
\draw[] (0,4) node[anchor=east] {$(0,2)$};
\draw[] (4,0) node[anchor=west] {$(2,0)$};
\draw[] (4,4) node[anchor=west] {$(2,2)$};
% Arrow labels
\draw[] (1,1) node[anchor=north] {$\tfrac{\partial}{\partial x}$};
\draw[] (3,3) node[anchor=north] {$\tfrac{\partial}{\partial x}$};
\draw[] (0,1) node[anchor=west] {$\tfrac{\partial}{\partial y}$};
\draw[] (0,3) node[anchor=west] {$\tfrac{\partial}{\partial y}$};
% The key and headings
\draw[black,thick,->] (-2.5,3.4)--(-2.5,4);
\draw[black,thick,->] (-2.5,3.4)--(-1.9,3.4);
\draw[] (-2.3,3.2) node[anchor=west] {$a$};
\draw[] (-2.7,3.6) node[anchor=south] {$b$};
\draw[] (2,4.6) node[anchor=south] {basis conversion};
\end{tikzpicture} 
\caption{The Laplace operator acting on vectors of $P_{n,k}=P_{n,k}^{(0,0)}$ coefficients has a sparse matrix representation if the range is represented as vectors of $\smash{P^{(2,2)}_{n,k}}$ coefficients. Here, the arrows indicate that the corresponding operation has a sparse matrix representation when the domain is $\smash{P_{n,k}^{(a,b)}}$ coefficients, where $(a,b)$ is at the tail of the arrow, and the range is $\smash{P_{n,k}^{(\tilde{a},\tilde{b})}}$ coefficients, where $(\tilde{a},\tilde{b})$ is at the head of the arrow. %For example, $\smash{\frac{\partial}{\partial x}}$ has a sparse matrix representation when the domain is represented in $\smash{P_{n,k}^{(0,0,0)}}$ (resp.~$\smash{P_{n,k}^{(1,0,1)}}$) and the range in $\smash{P_{n,k}^{(1,0,1)}}$ (resp.~$\smash{P_{n,k}^{(2,0,2)}}$).
}
\label{fig:Laplace} 
\end{figure}

>>>>

\documentclass{imposter}
%\documentclass[landscape]{imposter}

%\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
%\geometry{landscape}                		% Activate for rotated page geometry
\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
%\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex

\usepackage{wrapfig}								
\usepackage[font=scriptsize,labelfont=bf]{caption}
\usepackage[font=scriptsize]{subcaption}
\usepackage{float}
% \graphicspath{ {images/poster-eps/} }
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{bbm}

\usepackage{todonotes}

\graphicspath{ {../papers/halfdisk/images/} }
\graphicspath{ {images/} }

@@

\usepackage{natbib}
\usepackage{url}
\bibliographystyle{abbrv}

@@

\usepackage{listings}

\title{Sparse Spectral Methods on Disk Slices}
\author{Ben Snowball (Supervisors: Sheehan Olver (Main), Colin Cotter)}
\department{Department of Mathematics, MPE CDT}
\institution{Imperial College London}

%% Primary university logo.
%\logo{imperial.eps}
% Secondary logos (not supported in all poster styles)
%\logoa{AMCG.eps}

%% Options to set colors of multiple poster features.
\titlecolor{named}{Blue}
\postercolor{named}{Blue}
%\pagecolor{green}

\posterstyle{mpecdtstyle}

% This sets up boxes to have nice rounded corners and somewhat thicker lines
% than the default.
\psset{linecolor=postercolor,linewidth=.2,cornersize=absolute,linearc=0.5,framesep=0.5}

\begin{document}
%\showgrid

% LHS

  \leftbox{96}{
        \section{Abstract}

In recent years, sparse spectral methods have been derived using hierarchies of classical orthogonal polynomials (OPs) on intervals, disks, and triangles (e.g. \cite{olver2018recurrence}). Here, we extend this methodology to a hierarchy of non-classical orthogonal polynomials on disk slices (e.g. a half disk), building on the observation that sparsity is guaranteed due to the boundary being defined by an algebraic curve, and that the entries of partial differential operators can be determined using formulae in terms of (non-classical) univariate OPs.

    }

  \leftbox{81}{
      \section{Orthogonal polynomials on the half disk}

@@

Taking $\Omega$ as the half disk, the weight $W^{(a,b)}(x,y)$ results from setting:
\begin{align}
\begin{cases}
(\alpha,\beta, \gamma,\delta) &:= (0.1, -1, 1) \\
\rho(x) &:= (1-x^2)^{\half} \\
w_1(x) &:= x^a \: (1-x^2)^b \\
w_2(x) &:= (1-x^2)^b = (1-x)^b \: (1+x)^b.
\end{cases}
\end{align}

The weight $w_2(x)$ corresponds to the weight for the Jacobi polynomials \(\{P_n^{(b, b)}\}\). The weight $w_1(x)$ is non-classical and we introduce the notation \(\{H_n^{(a, b)}\}\) for the orthonormal polynomials. Thus we arrive at a two parameter family of 2D OPs \(\{\Pnk\}\) given by, for \(0 \le k \le n, \: n = 0,1,2,\dots,\)
\begin{align}
 \Pnk^{(a,b)}(x,y) := H_{n-k}^{(a, b+k+\half)}(x) \: \rho(x)^k \: P_k^{(b,b)}(\frac{y}{\rho(x)}), \quad (x,y) \in \Omega, 
\end{align}
orthogonal on \(\Omega\) with respect to the weight $W^{(a,b)}(x,y)$.

    }

    \leftbox{37}{
    	\section{Jacobi matrices and building the OPs}
Define, for some $N \in \No$:
\begin{align}
\bigPab_n := \begin{bmatrix}
		P^{(a,b)}_{n,-n}(x,y) \\
		\vdots \\
		P^{(a,b)}_{n,n}(x,y)
	\end{bmatrix} \in \R^{n+1}, 
\quad \quad 
\bigPab := \begin{bmatrix}
		\bigPab_0 \\
		\bigPab_1 \\
		\vdots \\
		\bigPab_N
	\end{bmatrix}.
\end{align}

\begin{lemma}
The three-term recurrences that $\Pnk^{(a,b)}(x,y)$ satisfies lead to Jacobi operators that correspond to multiplication by $x$ and $y$. Set $J_x^{(a,b)}, J_y^{(a,b)} \in \R^{\half (N+1)(N+2) \times \half (N+1)(N+2)}$ as the Jacobi matrices corresponding to
\begin{align}
J_x^{(a,b)} \bigPab(x,y) = x \bigPab(x,y), \quad J_y^{(a,b)} \bigPab(x,y) = y \bigPab(x,y).
\end{align}
Then  \(J_x^{(a,b)}, J_y^{(a,b)}\) have similar sparse (block-tridiagonal) Banded-Block-Banded structures.
\end{lemma}
Combining each system in (4) we can write this as a block-tridiagonal system of equations. Applying a preconditioner matrix to the system \citep{dunkl2014orthogonal} (p78), we can expand the resulting lower-triangular system to obtain the recurrence:
    \begin{align}
\begin{cases}
\bigPab_{-1}(x,y) := 0, \quad \bigPab_{0}(x,y) := P^{(a,b)}_{0,0} \\
\bigPab_{n+1}(x,y) = -\Dnt (B_n-G_n(x,y)) \bigPab_n(x,y) - \Dnt C_n  \,\bigPab_{n-1}(x,y).
\end{cases}
\end{align}
where the matrix coefficients $\Dnt, B_n, C_n, G_n(x,y)$ can be explicitly found.
    }

    \rightbox{96}{
    	\section{Sparse partial differential operators}

For general \(a,b\) and general functions \(f(x,y) = \bigPab(x,y)^\top \mathbf{f}\), there exist sparse operators \(D_x^{(a,b)}, \: D_y^{(a,b)}, \: W_x^{(a,b)}, \: W_y^{(a,b)}\) such that (as seen in the DLMF \cite{DLMFDerivatives} for other OP families):
\begin{align*}
\dx[f(x,y)] &= \bigP^{(a+1,b+1)}(x,y)^T \: D_x^{(a,b)} \: \mathbf{f}, \\
\dy[f(x,y)] &= \bigP^{(a,b+1)}(x,y)^T \: D_y^{(a,b)} \: \mathbf{f}, \\
\dx[\Wab(x,y) \: f(x,y)] &= W^{(a-1,b-1)}(x,y) \: \bigP^{(a-1,b-1)}(x,y)^T \: W_x^{(a,b)} \: \mathbf{f}, \\
\dy[\Wab(x,y) \: f(x,y)] &= W^{(a,b-1)}(x,y) \: \bigP^{(a,b-1)}(x,y)^T \: W_y^{(a,b)} \: \mathbf{f}.
\end{align*}
Further, we note that there exist Banded-Block-Banded matrix operators $T^{(1,0)\to(0,0)}\), \(T^{(0,1)\to(1,1)}$ such that:
\begin{align*}
x \bigP^{(1,0)}(x,y) &= \Big(T^{(1,0)\to(0,0)} \Big)^T \: \bigP^{(0,0)}(x,y) \\
\bigP^{(0,1)}(x,y) &= \Big(T^{(0,1)\to(1,1)} \Big)^T \: \bigP^{(1,1)}(x,y)
\end{align*}
\begin{align}
    \implies \Delta^{(1,1)} := D_x^{(0,0)} \: W_x^{(1,1)} + T^{(0,1)\to(1,1)} \: D_y^{(0,0)} \: T^{(1,0)\to(0,0)} \: W_y^{(1,1)},
\end{align}

\begin{wrapfigure}{R}{0.7\textwidth}
        \center 
   \begin{subfigure}[t]{0.35\textwidth}
         \centering
        \includegraphics[scale=1.0]{sparsityoflaplacian}
        \label{fig:sparsity}
    \end{subfigure}
   \begin{subfigure}[t]{0.3\textwidth}
         \centering
        \includegraphics[scale=0.6]{solution}
        \label{fig:solution}
    \end{subfigure}
    \begin{subfigure}[t]{0.3\textwidth}
                \center
        \begin{tikzpicture} 
        % Box/square (4x4)
        \draw[black,solid,ultra thin] (0,0)--(0,4);
        \draw[black,solid,ultra thin] (0,0)--(4,0);
        \draw[black,solid,ultra thin] (4,0)--(4,4);
        \draw[black,solid,ultra thin] (0,4)--(4,4);
        % Dots
        \draw[black,fill=black] (0,0) circle (.5ex);
        \draw[black,fill=black] (2,2) circle (.5ex);
        \draw[black,fill=black] (4,4) circle (.5ex);
        \draw[black,fill=black] (0,2) circle (.5ex);
        \draw[black,fill=black] (0,4) circle (.5ex);
        % Arrows
        \draw[black,thick,->] (0.2,0.2)--(1.8,1.8);
        \draw[black,thick,->] (2.2,2.2)--(3.8,3.8);
        \draw[black,thick,->] (0,0.2)--(0,1.8);
        \draw[black,thick,->] (0,2.2)--(0,3.8);
        \draw [black,thick,->] (0.2,4.2) to [out=20,in=160] (3.8,4.2);
        % Node (parameter) labels
        \draw[] (0,0) node[anchor=east] {$(0,0)$};
        \draw[] (0,4) node[anchor=east] {$(0,2)$};
        \draw[] (4,0) node[anchor=west] {$(2,0)$};
        \draw[] (4,4) node[anchor=west] {$(2,2)$};
        % Arrow labels
        \draw[] (1,1) node[anchor=north] {$\tfrac{\partial}{\partial x}$};
        \draw[] (3,3) node[anchor=north] {$\tfrac{\partial}{\partial x}$};
        \draw[] (0,1) node[anchor=west] {$\tfrac{\partial}{\partial y}$};
        \draw[] (0,3) node[anchor=west] {$\tfrac{\partial}{\partial y}$};
        % The key and headings
        \draw[black,thick,->] (-2.5,3.4)--(-2.5,4);
        \draw[black,thick,->] (-2.5,3.4)--(-1.9,3.4);
        \draw[] (-2.3,3.2) node[anchor=west] {$a$};
        \draw[] (-2.7,3.6) node[anchor=south] {$b$};
        \draw[] (2,4.6) node[anchor=south] {basis conversion};
        \end{tikzpicture} 
	\label{fig:Laplace}
   \end{subfigure}
    \centering
    \captionsetup{width=.7\linewidth}
    \caption{Left: A "spy" plot of the $\Delta$ operator for $\Wii(x,y) \: \bigPii(x,y) \to \bigPii(x,y)$ showing its sparsity. Right: The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 ? 10((x ? 0.5)^2 + y^2)))$. Bottom: Diagram illustrating how the parameters change under partial differentiation. }
\end{wrapfigure}

where $\Delta^{(1,1)}$ is the matrix operator for the Laplacian, that will take us from coefficients for expansion in the weighted space $\Wii(x,y) \: \bigPii(x,y)$ to coefficients in the non-weighted space $\bigPii(x,y)$. This operator will have Banded-Block-Banded structure, and hence sparse. Note that this construction ensures imposition of Dirichlet zero boundary conditions.
